\vspace{-3em} \LARGE Hamiltonian Privilege

### 2023-02-13 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
ClassicalMechanicsPhil: adding comments for referee

\noindent
This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system. We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
% % % For further discussion, see \textcites{shpmp}. [omit this for review].

@@

When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold. For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$. \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map. To avoid confusion, we will call the coordinates of the state space $\mathcal{S}$ ``state variables.'' As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$. We will neglect these subtleties in what follows.} The map $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$. To represent the distribution $\rho_c$ in terms of state variables, we define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ by $\rho_{c, \xi} \equiv (\rho_c \circ \xi^{-1})$. This density plays a crucial role throughout our argument. It allows us to compare and track the amount of matter in different possible states.

@@

\noindent In our case, this applies to both the state of the whole system and the state of its parts.

@@

\noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.

@@

In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^j$. Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_k}$ must be invertible. Let $H$ denote the Hamiltonian. By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, so the Hessian $\frac{\partial^2 H}{\partial p_k \partial p_i} = \frac{\partial}{\partial p_k} (\frac{\partial H}{\partial p_i})$ equals the Jacobian $\frac{\partial v^i}{\partial p_k}$. Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive. Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables. These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle). Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian. As discussed in Section~\ref{Curiel}, Lagrangian mechanics only yields a unique evolution in the hyper-regular domain.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry. As we discuss in Section~\ref{Curiel}, Equation~\ref{weak_equivalence} corresponds to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
%; otherwise, the positions and velocities do not determine how the system evolves

>>>>

\noindent
This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system. We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} The general idea is that this assumption tells us that the state of a system is captured by a statistical distribution over the state space of the parts. If the state space of the parts is a manifold, this must be differentiable so that it can support a density at each point. Moreover, given that this density must be a function of the state, and not of the coordinates, we find that the manifold must be symplectic.
% % % For further discussion, see \textcites{shpmp}. [omit this for review].

@@

When the system is characterized by continuous quantities, $\mathcal{S}$ can be taken to be a manifold.\footnote{The use of manifolds is, technically, a choice of how the physics is represented. One could make another choice, for example, by representing the state variables not as local coordinates but as an algebra of operators that act on the state. In this case, we believe one should be able to run a similar argument and find the analogue of a symplectic structure for the different mathematical representation.} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$. \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map. To avoid confusion, we will call the coordinates of the state space $\mathcal{S}$ ``state variables.'' As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$. We will neglect these subtleties in what follows.} The map $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$. To represent the distribution $\rho_c$ in terms of state variables, we define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ by $\rho_{c, \xi} \equiv (\rho_c \circ \xi^{-1})$. This density plays a crucial role throughout our argument. It allows us to compare and track the amount of matter in different possible states.

@@

\noindent In our case, this applies to both the state of the whole system and the state of its parts. Under this assumption we will find the symplectic structure must be preserved in time, which gives us Hamiltonian evolution.

@@

\noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space. This assumption means that momentum can be expressed as a function of position and velocity, which mathematically corresponds to the case in which the Hamiltonian is hyper-regular. In this regime we find Lagrangian mechanics. Moreover, to be able to express the symplectic form, and therefore the density, over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be the one of a massive particle under scalar and vector potential forces.

@@

In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^j$. Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_k}$ must be invertible. Let $H$ denote the Hamiltonian. By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, so the Hessian $\frac{\partial^2 H}{\partial p_k \partial p_i} = \frac{\partial}{\partial p_k} (\frac{\partial H}{\partial p_i})$ equals the Jacobian $\frac{\partial v^i}{\partial p_k}$. Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive. This, by definition, means that motion equivalence implies the Hamiltonian is hyper-regular. Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables. These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle). Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian. As discussed in Section~\ref{Curiel}, Lagrangian mechanics only yields a unique evolution in the hyper-regular domain.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry. As we discuss in Section~\ref{Curiel}, Equation~\ref{weak_equivalence} corresponds to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
%; otherwise, the positions and velocities do not determine how the system evolves
